---
title: 'Chapter 1: k-Nearest Neighbors (kNN)'
author: "by Edneide Ramalho"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
      highlight: textmate
      logo: logo.png
      theme: flatly
      number_sections: yes
      toc: yes
      toc_float:
        collapsed: yes
        smooth_scroll: no
---

# Libraries

```{r}
library(class)
library(tidyverse)
```

# Data sets

```{r}
loans <- read.csv("https://assets.datacamp.com/production/repositories/718/datasets/7805fceacfb205470c0e8800d4ffc37c6944b30c/loans.csv")

locations <- read.csv("https://assets.datacamp.com/production/repositories/718/datasets/571628c39048df59c40c9dcfba146a2cf7a4a0e3/locations.csv")

knn_traffic_signs <- read.csv("https://assets.datacamp.com/production/repositories/718/datasets/c274ea22cc3d7e12d7bb9fdc9c2bdabe9ab025f4/knn_traffic_signs.csv")

donors <- read.csv("https://assets.datacamp.com/production/repositories/718/datasets/9055dac929e4515286728a2a5dae9f25f0e4eff6/donors.csv")


```

# Exercises

## **Recognizing a road sign with kNN**

After several trips with a human behind the wheel, it is time for the self-driving car to attempt the test course alone.

As it begins to drive away, its camera captures the following image:

Can you apply a kNN classifier to help the car recognize this sign?

The dataset `signs` is loaded in your workspace along with the dataframe `next_sign`, which holds the observation you want to classify.

-   Load the `class` package.

-   Create a vector of sign labels to use with kNN by extracting the column `sign_type` from `signs`.

-   Identify the `next_sign` using the `knn()` function.

    -   Set the `train` argument equal to the `signs` data frame *without* the first column.

    -   Set the `test` argument equal to the data frame `next_sign`.

    -   Use the vector of labels you created as the `cl` argument.

```{r}
signs <- knn_traffic_signs %>% 
  dplyr::select(sign_type:b16)

next_sign <- knn_traffic_signs %>% 
  dplyr::filter(sample == "example")
next_sign <- next_sign %>% 
  dplyr::select(r1:b16)
```

```{r}
# Load the 'class' package
library(class)

# Create a vector of labels
sign_types <- signs$sign_type

# Classify the next sign observed
knn(train = signs[-1], test = next_sign, cl = sign_types)
```

Based on the kNN algorithm, the next sign is a Stop sign.
