---
title: "Interpreting regression models"
author: "by Edneide Ramalho"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
      highlight: textmate
      logo: logo.png
      theme: flatly
      number_sections: yes
      toc: yes
      toc_float:
        collapsed: yes
        smooth_scroll: no
---

# Interpretation of the coefficients

A fitted model for the poverty rate of U.S. counties as a function of high school graduation rate is:

$$\hat{poverty} =  64.594 - 0.591 \cdot hs\_grad$$
The interpretation of the coefficients is the following:

Among U.S. counties, each additional percentage point increase in the high school graduation rate is associated with about a 0.591 percentage point decrease in the poverty rate.

# Fitting simple linear models

While the `geom_smooth(method = "lm")` function is useful for drawing linear models on a scatterplot, it doesn't actually return the characteristics of the model. As suggested by that syntax, however, the function that creates linear models is `lm()`. This function generally takes two arguments:

- A `formula` that specifies the model
- A `data` argument for the data frame that contains the data you want to use to fit the model

The `lm()` function return a model object having class `"lm"`. This object contains lots of information about your regression model, including the data used to fit the model, the specification of the model, the fitted values and residuals, etc.

- Using the bdims dataset, create a linear model for the weight of people as a function of their height.

- Using the mlbBat10 dataset, create a linear model for SLG as a function of OBP.
- Using the mammals dataset, create a linear model for the body weight of mammals as a function of their brain weight, after taking the natural log of both variables.

```{r}
library(openintro)
library(ggplot2)
library(tidyverse)
```

```{r}
# Linear model for weight as a function of height
glimpse(bdims)
lm(wgt ~ hgt, data = bdims)
```

```{r}
# Linear model for SLG as a function of OBP
glimpse(mlbbat10)
lm(slg ~ obp, data = mlbbat10)
```

```{r}
# Log-linear model for body weight as a function of brain weight
glimpse(mammals)
lm(log(body_wt) ~ log(brain_wt), data = mammals)
```


# Your linear model object