---
title: "Easystats Performance"
author: "by Edneide Ramalho"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
      highlight: textmate
      logo: logo.png
      theme: flatly
      number_sections: yes
      toc: yes
      toc_float:
        collapsed: yes
        smooth_scroll: no
---

This script is based on the video:

<https://www.youtube.com/watch?v=Bi8sHIo3s1Y>

from the Business Science YouTube Chanel

# Libraries

```{r}
# install.packages("performance", dependences = TRUE)
```

```{r}
library(tidyverse)
library(performance)    
```

# Data

```{r}
glimpse(mpg)
```

We are going to perform a linear regression and check the model performance.

# Performance

```{r}
model_lm <- lm(hwy ~ displ + class, data = mpg)

model_lm

check_model(model_lm)
```

-   In the first plot we check the **linearity**, the reference line should be flat and horizontal. The homogeneity of variance plot give us a similar result, where the reference line should be flat and horizontal as well.

-   For the **collinearity plot**: High bars indicate collinearity issues. Features with high collinearity should be removed from models that don't include regularization. (e. g. Linear Regression).

-   For the **Influential Observations plot**: We check if there is any point outside of the dashed lines., you may want to remove this point of your model to improve performance.

-   **Q-Q Plot**: See how model residuals compare to a Normal Distribution. the points should follow the line. In the plot, we can see that in the end, points fall from the green line. We can investigate those points.

-   **Normality Residuals**: make sure the faded blue distribution is close to the green normal curve. We see some skew towards the right.

# Tidymodels

-   Tidymodels are similar to Scikit Learn in Python.

-   A modeling and machine learning framework in R.

```{r}
library(tidymodels)
```

## Linear regression

```{r}
model_lm_tidy <- linear_reg() %>% 
  set_engine("lm") %>% 
  fit(hwy ~ displ + class, data = mpg)
```

```{r}
check_model(model_lm_tidy)
```
