---
title: "Wittgenstein Centre Human Capital Data Explorer"
author: "by Edneide Ramalho"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: 
    html_document:
      highlight: textmate
      logo: logo.png
      theme: flatly
      number_sections: yes
      toc: yes
      toc_float:
        collapsed: yes
        smooth_scroll: no
---

# Installation

```{r}
# install.packages("wcde")
```

# Example

```{r}
library(wcde)
```

## Data

-   SSP2 education specific tfr for Austria:

```{r}
get_wcde(indicator = "etfr", country_name = "Austria")
```

-   SSP2 education specific population sizes for Iran and Kenya:

```{r}
get_wcde(indicator = "epop", country_code = c(364, 404))
```

-   SSP1, 2 and 3 gender gaps in educational attainment (15+) for all countries:

```{r}
get_wcde(indicator = "ggapedu15", scenario = 1:3)
```

# Population pyramids

Create population pyramids by setting male population values to negative equivalent to allow for divergent columns from de y axis.

```{r}
library(tidyverse)
d <- get_wcde(indicator = "epop", country_code = 900)
sample_n(d, 5)
```

```{r}
w <- d %>% 
  edu_group_sum(n = 4) %>% 
  mutate(pop = ifelse(test = sex == "Male", yes = -epop, no = epop),
         pop = pop/1e3)
sample_n(w, 4)
```

## Standard plot

```{r}
library(lemon)

w %>%
  filter(year == 2020) %>%
  ggplot(mapping = aes(x = pop, y = age, fill = fct_rev(education))) +
  geom_col() +
  geom_vline(xintercept = 0, colour = "black") + 
  scale_x_symmetric(labels = abs) +
  scale_fill_manual(values = wic_col4, name = "Education") +
  labs(x = "Population (millions)", y = "Age") +
  theme_bw() 
```

```{r}
w <- w %>% 
  dplyr::mutate(pop_max = ifelse(sex == "Male", -max(pop), max(pop)))

w %>% 
  filter(year == 2020) %>%
  ggplot(mapping = aes(x = pop, y = age, fill = fct_rev(education))) +
  geom_col() +
  geom_vline(xintercept = 0, colour = "black") +
  scale_x_continuous(labels = abs, expand = c(0, 0)) +
  scale_fill_manual(values = wic_col4, name = "Education") +
  labs(x = "Population (millions)", y = "Age") +
  facet_wrap(facets = "sex", scales = "free_x", strip.position = "bottom") +
  geom_blank(mapping = aes(x = pop_max * 1.1)) +
  theme(panel.spacing.x = unit(0, "pt"),
        strip.placement = "outside",
        strip.background = element_rect(fill = "transparent"),
        strip.text.x = element_text(margin = margin( b = 0, t = 0)))
```

# More information

For more information go to the package website: <https://guyabel.github.io/wcde/index.html>
