---
title: "Chapter 3 - Full, Semi and Anti Joins"
author: "by Edneide Ramalho"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
      highlight: textmate
      logo: logo.png
      theme: readable
      number_sections: yes
      toc: yes
      toc_float:
        collapsed: yes
        smooth_scroll: no
---

# Full join

It keeps all of both tables.

-   Importing the data frames:

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
inventories <- read_csv("dados/inventories.csv")
inventory_parts <- read_csv("dados/inventory_parts.csv")

inventory_parts_joined <- inventories %>% 
  inner_join(inventory_parts, by = c("id" = "inventory_id")) %>% 
  select(-id, -version) %>% 
  arrange(desc(quantity))

batmobile <- inventory_parts_joined %>% 
  filter(set_num == "7784-1") %>% 
  select(-set_num)

batwing <- inventory_parts_joined %>% 
  filter(set_num == "70916-1") %>% 
  select(-set_num)
```

```{r message=FALSE, warning=FALSE}
batmobile %>% 
  full_join(batwing, by = c("part_num", "color_id"), suffix = c("_batmobile", "_batwing"))
```

## Replace NA: multiple variables

```{r message=FALSE, warning=FALSE}
batmobile %>% 
  full_join(batwing, by = c("part_num", "color_id"), suffix = c("_batmobile", "_batwing")) %>% 
  replace_na(list(quantity_batmobile = 0,
                  quantity_batwing = 0))
```

## Exercises

### Differences between Batman and Star Wars

In the video, you compared two sets. Now, you'll compare two themes, each of which is made up of many sets.

First, you'll need to join in the `themes`. Recall that doing so requires going through the `sets` first. You'll use the `inventory_parts_joined` table from the video, which is already available to you in the console.

    inventory_parts_joined <- inventories %>%
      inner_join(inventory_parts, by = c("id" = "inventory_id")) %>%
      arrange(desc(quantity)) %>%
      select(-id, -version)

-   In order to join in the themes, you'll first need to combine the `inventory_parts_joined` and `sets` tables. Keep only the observations that have a match in both tables.

-   Then, combine the first join with the `themes` table, using the suffix argument to clarify which table each `name` came from (`"_set"` or `"_theme"`). Keep only the observations that have a match in both tables.

*If your exercise is timing out, you are likely joining the wrong tables, causing the join result to be too big! Check your code carefully!*

```{r}
library(readr)
sets <- read_csv("dados/sets.csv")
themes <- read_csv("dados/themes.csv")
```

```{r}
# Start with inventory_parts_joined table
inventory_parts_joined %>%
  # Combine with the sets table 
  inner_join(sets, by = "set_num") %>% 
  # Combine with the themes table 
  inner_join(themes, by = c("theme_id" = "id"),
            suffix = c("_set", "_theme")) 
```

### **Aggregating each theme**

Previously, you combined tables to compare themes. Before doing this comparison, you'll want to aggregate the data to learn more about the pieces that are a part of each theme, as well as the colors of those pieces.

The table you created previously has been preloaded for you as `inventory_sets_themes`. It was filtered for each theme, and the objects have been saved as `batman` and `star_wars`.

```{r}
inventory_sets_themes <- inventory_parts_joined %>%
  inner_join(sets, by = "set_num") %>%
  inner_join(themes, by = c("theme_id" = "id"), suffix = c("_set", "_theme"))

batman <- inventory_sets_themes %>%
  filter(name_theme == "Batman")

star_wars <- inventory_sets_themes %>%
  filter(name_theme == "Star Wars")
```

-   Count the part number and color id for the parts in Batman and Star Wars, weighted by quantity.

```{r}
batman %>% 
  count(part_num,color_id, wt = quantity)
```

```{r}
star_wars %>% 
  count(part_num, color_id, wt = quantity)
```

### **Full-joining Batman and Star Wars LEGO parts**

Now that you've got separate tables for the pieces in the `batman` and `star_wars` themes, you'll want to be able to combine them to see any similarities or differences between the two themes. The aggregating from the last exercise has been saved as `batman_parts` and `star_wars_parts`, and is preloaded for you.

```{r}
batman_parts <- batman %>%
  count(part_num, color_id, wt = quantity)

star_wars_parts <- star_wars %>%
  count(part_num, color_id, wt = quantity)
```

-   Combine the `star_wars_parts` table with the `batman_parts` table; use the suffix argument to include the `"_batman"` and `"_star_wars"` suffixes.

-   Replace all the NA values in the `n_batman` and `n_star_wars` columns with 0s.

```{r}
batman_parts %>%
  # Combine the star_wars_parts table 
  full_join(star_wars_parts, by = c("part_num", "color_id"),
            suffix = c("_batman", "_star_wars")) %>% 
  # Replace NAs with 0s in the n_batman and n_star_wars columns 
  replace_na(list(n_batman = 0,
                  n_star_wars = 0))

```

### **Comparing Batman and Star Wars LEGO parts**

The table you created in the last exercise includes the part number of each piece, the color id, and the number of each piece in the Star Wars and Batman themes. However, we have more information about each of these parts that we can gain by combining this table with some of the information we have in other tables. Before we compare the themes, let's ensure that we have enough information to make our findings more interpretable. The table from the last exercise has been saved as `parts_joined` and is preloaded for you.

```{r}
parts_joined <- batman_parts %>%
  full_join(star_wars_parts, by = c("part_num", "color_id"), suffix = c("_batman", "_star_wars")) %>%
  replace_na(list(n_batman = 0, n_star_wars = 0))

```

-   Sort the number of star wars pieces in the `parts_joined` table in descending order.

-   Join the `colors` table to the `parts_joined` table.

-   Combine the `parts` table to the previous join; add `"_color"` and `"_part"` suffixes to specify whether or not the information came from the `colors` table or the `parts` table.

```{r}
# Importing colors table
library(readr)
colors <- read_csv("dados/colors.csv")
glimpse(colors)
```

```{r}
# Importing parts table
library(readr)
parts <- read_csv("dados/parts.csv")
glimpse(parts)
```

```{r}
parts_joined %>% 
  # Join the number of star wars pieces in descending order
  dplyr::arrange(desc(n_star_wars)) %>% 
  # Join the colors table to the parts_joined table
  inner_join(colors, by = c("color_id" = "id")) %>% 
  # Join the parts table to the previous join
  inner_join(parts, by = "part_num", suffix = c("_color", "_part"))
```

# The semi- and anti-join verbs

## Filtering joins

-   Keeps or removes observations from the first table
-   Doesn't add new variables
-   `semi_join()` What observations in X are **also** in Y?
-   `anti_join()` What observations in X are not in Y?

## The semi join

```{r}
glimpse(batmobile)
```

```{r}
glimpse(batwing)
```

```{r}
batmobile %>% 
  semi_join(batwing, by = c("color_id", "part_num"))
```

When we do a semi_join we reduce the data frame. Only color_ids and part_num in batmobile that are also in Batwing will be returned.

## The anti join

It is the opposite of semi join. We are asking what observations in batmobile are not in batwing?

```{r}
batmobile %>% 
  anti_join(batwing, by = c("color_id", "part_num"))
```

## The joining verbs

-   `inner_join`

-   `left_join`

-   `right_join`

-   `full_join` -\> Keep all observations from both tables.

-   `semi_join` -\> Filter the first table for observations that match the second.

-   `anti_join` -\> Filter the first table for observations that don't match the second.

## Exercises

### **Something within one set but not another**

In the videos, you learned how to filter using the semi- and anti-join verbs to answer questions you have about your data. Let's focus on the `batwing` dataset, and use our skills to determine which parts are in both the `batwing` and `batmobile` sets, and which sets are in one, but not the other. While answering these questions, we'll also be determining whether or not the parts we're looking at in both sets also have the same color in common.

-   Filter the batwing set for parts that are **also** in the batmobile, whether or not they have the same color.

```{r}
batwing %>% 
  semi_join(batmobile, by = "part_num")
```

-   Filter the batwing set for parts that **aren't** also in the batmobile, whether or not they have the same color.

    ```{r}
    batwing %>% 
      anti_join(batmobile, by = "part_num")
    ```

### What colors are included in at least one set?

Besides comparing two sets directly, you could also use a filtering join like `semi_join` to find out which colors ever appear in any inventory part. Some of the colors could be optional, meaning they aren't included in any sets.

The `inventory_parts` and `colors` tables have been preloaded for you.

-   Use the `inventory_parts` table to find the colors that are included in at least one set.

```{r}
# Use inventory_parts to find colors included in at least one set
colors %>% 
    semi_join(inventory_parts, by = c("id" = "color_id"))
```

### Which set is missing version 1?

Each set included in the LEGO data has an associated version number. We want to understand the version we are looking at to learn more about the parts that are included. Before doing that, we should confirm that there aren't any sets that are missing a particular version.

Let's start by looking at the first version of each set to see if there are any sets that don't include a first version.

-   Use `filter()` to extract `version` `1` from the `inventories` table; save the filter to `version_1_inventories`.

-   Use `anti_join` to combine `version_1_inventories` with `sets` to determine which set is missing a version 1.

```{r}
# Use filter() to extract version 1 
version_1_inventories <- inventories %>%
    filter(version == 1)


# Use anti_join() to find which set is missing a version 1
sets %>% 
    anti_join(version_1_inventories, by = "set_num")
```

# Visualizing set differences

## Aggregating sets into colors

```{r}
batmobile_colors <- batmobile %>% 
  group_by(color_id) %>% 
  summarize(total = sum(quantity))

batmobile_colors
```

```{r}
batwing_colors <- batwing %>% 
  group_by(color_id) %>% 
  summarize(total = sum(quantity))

batwing_colors
```

## Comparing color schemes of sets

```{r}
colors_df <- batmobile_colors %>% 
  full_join(batwing_colors, by = "color_id", 
            suffix = c("_batmobile", "_batwing")) %>% 
  replace_na(list(total_batmobile = 0, 
                  total_batwing = 0))
colors_df
```

## Adding the colors names

```{r}
colors_df <- colors_df %>% 
  inner_join(colors, by = c("color_id" = "id"))
colors_df
```

## Adding percentages

```{r}
colors_df <- colors_df %>% 
  mutate(total_batmobile = total_batmobile/sum(total_batmobile),
         total_batwing = total_batwing/sum(total_batwing))
colors_df
```

## The difference between fractions

```{r}
colors_joined <- colors_df %>% 
  mutate(difference = total_batmobile - total_batwing)
colors_joined
```

## Visualizing the data

```{r}
library(ggplot2)
library(forcats)

color_palette <- setNames(paste0("#", colors_joined$rgb), colors_joined$name)

colors_joined %>% 
  mutate(name = fct_reorder(name, difference)) %>% 
  ggplot(aes(name, difference, fill = name)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values = color_palette, guide = FALSE)
```

Colors with positive difference are more common in the batmobile set, while colors with negative difference are more common in the batwing set.

## Exercises

### **Aggregating sets to look at their differences**

To compare two individual sets, and the kinds of LEGO pieces that comprise them, we'll need to aggregate the data into separate themes. Additionally, as we saw in the video, we'll want to add a column so that we can understand the percentages of specific pieces that are part of each set, rather than looking at the numbers of pieces alone.

The `inventory_parts_themes` table has been preloaded for you.

```{r}
inventory_parts_themes <- inventories %>%
  inner_join(inventory_parts, by = c("id" = "inventory_id")) %>%
  arrange(desc(quantity)) %>%
  select(-id, -version) %>%
  inner_join(sets, by = "set_num") %>%
  inner_join(themes, by = c("theme_id" = "id"), suffix = c("_set", "_theme"))
```

-   Add a filter for the `"Batman"` set to create the `batman_colors` object.

-   Add a percent column to `batman_colors` that displays the total divided by the sum of the total.

-   Repeat the steps to filter and aggregate the `"Star Wars"` set data to create the `star_wars_colors` object.

    -   Again, add a percent column to display the percent of the total.

```{r}
batman_colors <- inventory_parts_themes %>%
  # Filter the inventory_parts_themes table for the Batman theme
  filter(name_theme == "Batman") %>%
  group_by(color_id) %>%
  summarize(total = sum(quantity)) %>%
  # Add a percent column of the total divided by the sum of the total 
  mutate(percent = total/sum(total))

# Filter and aggregate the Star Wars set data; add a percent column
star_wars_colors <- inventory_parts_themes %>%
  filter(name_theme == "Star Wars") %>% 
  group_by(color_id) %>% 
  summarize(total = sum(quantity)) %>% 
  mutate(percent = total/sum(total))
```

### **Combining sets**

The data you aggregated in the last exercise has been preloaded for you as `batman_colors` and `star_wars_colors`. Prior to visualizing the data, you'll want to combine these tables to be able to directly compare the themes' colors.

    batman_colors <- inventory_parts_themes %>%
      filter(name_theme == "Batman") %>%
      group_by(color_id) %>%
      summarize(total = sum(quantity)) %>%
      mutate(percent = total / sum(total))

    star_wars_colors <- inventory_parts_themes %>%
      filter(name_theme == "Star Wars") %>%
      group_by(color_id) %>%
      summarize(total = sum(quantity)) %>%
      mutate(percent = total / sum(total))

-   Join the `batman_colors` and `star_wars_colors` tables; be sure to include all observations from both tables.

-   Replace the NAs in the `total_batman` and `total_star_wars` columns.

```{r}
batman_colors %>%
  # Join the Batman and Star Wars colors
  full_join(star_wars_colors, by = "color_id", suffix = c("_batman", "_star_wars")) %>%
  # Replace NAs in the total_batman and total_star_wars columns
  replace_na(list(total_batman = 0, 
                  total_star_wars = 0)) %>%
  inner_join(colors, by = c("color_id" = "id")) 
```

-   Add a `difference` column that calculates the difference between `percent_batman` and `percent_star_wars`, and a `total` column, which is the sum of `total_batman` and `total_star_wars`.

-   Add a `filter` to select observations where `total` is at least 200.

```{r}
batman_colors %>%
  full_join(star_wars_colors, by = "color_id", suffix = c("_batman", "_star_wars")) %>%
  replace_na(list(total_batman = 0, total_star_wars = 0)) %>%
  inner_join(colors, by = c("color_id" = "id")) %>%
  # Create the difference and total columns
  mutate(difference = percent_batman - percent_star_wars,
         total = total_batman + total_star_wars) %>%
  # Filter for totals greater than 200
  filter(total >= 200)
```

### **Visualizing the difference: Batman and Star Wars**

In the last exercise, you created `colors_joined`. Now you'll create a bar plot with one bar for each color (`name`), showing the difference in percentages.

Because factors and visualization are beyond the scope of this course, we've done some processing for you: here is the code that created the `colors_joined` table that will be used in the video.

```{r}
colors_joined <- batman_colors %>%
  full_join(star_wars_colors, by = "color_id", suffix = c("_batman", "_star_wars")) %>%
  replace_na(list(total_batman = 0, total_star_wars = 0)) %>%
  inner_join(colors, by = c("color_id" = "id")) %>%
  mutate(difference = percent_batman - percent_star_wars,
         total = total_batman + total_star_wars) %>%
  filter(total >= 200) %>%
  mutate(name = fct_reorder(name, difference)) 
```

```{r}
colors_joined <- colors_joined %>% 
  mutate(rgb = paste0("#", rgb))
```

-   Create a bar plot using the `colors_joined` table to display the most prominent colors in the Batman and Star Wars themes, with the bars colored by their `name`.

```{r}
color_palette <- setNames(colors_joined$rgb, colors_joined$name)
# Create a bar plot using colors_joined and the name and difference columns
ggplot(colors_joined, aes(name, difference, fill = name)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values = color_palette, guide = FALSE) +
  labs(y = "Difference: Batman - Star Wars")
```
