---
title: "Random Numbers and Probability"
author: "by Edneide Ramalho"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: 
    html_document:
      highlight: textmate
      logo: logo.png
      theme: flatly
      number_sections: yes
      toc: yes
      toc_float:
        collapsed: yes
        smooth_scroll: no
---


# Sampling from a data frame

```{r}
library(dplyr)
name <- c("Amir", "Brian", "Claire", "Damian")
n_sales <- c(178, 126, 75, 69)
sales_counts <- tibble::tibble(name = name,
                               n_sales = n_sales)
```

```{r}
set.seed(5)
sales_counts %>% 
  sample_n(1)
```

```{r}
set.seed(5)
sales_counts %>% 
  sample_n(2)
```

```{r}
sales_counts %>% 
  sample_n(2, replace = TRUE)
```

```{r}
sample(sales_counts, 5, replace = TRUE)
```

```{r}
library(readr)
seller_1 <- read_csv("data/seller_1.csv")
# Count the dealsfor each product
amir_deals <- seller_1
amir_deals %>%
  count(product)
```

-   Create a new column called `prob` by dividing `n` by the total number of deals Amir worked on.

```{r}
amir_deals %>% 
  count(product) %>% 
  mutate(prob = n/sum(n))
```

-   Set the random seed to `31`.

-   Take a sample of 5 deals **without** replacement.

```{r}
# Set random seed to 31
set.seed(31)

# Sample 5 deals without replacement
amir_deals %>%
  sample_n(5, replace = FALSE)
```

-   Take a sample of 5 deals **with** replacement.

```{r}
# Set random seed to 31
set.seed(31)

# Sample 5 deals with replacement
amir_deals %>%
  sample_n(5, replace = TRUE)

```

# Discrete distributions

## Creating a probability distribution 

A new restaurant opened a few months ago, and the restaurant's management wants to optimize its seating space based on the size of the groups that come most often. On one night, there are 10 groups of people waiting to be seated at the restaurant, but instead of being called in the order they arrived, they will be called randomly. In this exercise, you'll investigate the probability of groups of different sizes getting picked first. Data on each of the ten groups is contained in the restaurant_groups data frame.

Remember that expected value can be calculated by multiplying each possible outcome with its corresponding probability and taking the sum. The restaurant_groups data is available and dplyr and ggplot2 are loaded.

```{r}
library(readr)
restaurant_groups <- read_csv("restaurant_groups.csv")
```

- Create a histogram of the group_size column of restaurant_groups, setting the number of bins to 5.

```{r}
library(ggplot2)
ggplot(restaurant_groups, aes(x = group_size)) +
  geom_histogram(bins = 5)
```

Count the number of each group_size in restaurant_groups, then add a column called probability that contains the probability of randomly selecting a group of each size. Store this in a new data frame called size_distribution.

```{r}
# Create probability distribution
size_distribution <- restaurant_groups %>%
  # Count number of each group size
  count(group_size) %>%
  # Calculate probability
  mutate(probability = n / sum(n))

size_distribution
```

Calculate the expected value of the size_distribution, which represents the expected group size.

```{r}
# Create probability distribution
size_distribution <- restaurant_groups %>%
  count(group_size) %>%
  mutate(probability = n / sum(n))

# Calculate expected group size
expected_val <- sum(size_distribution$group_size *
                    size_distribution$probability)
expected_val
```

Calculate the probability of randomly picking a group of 4 or more people by filtering and summarizing.

```{r}
# Create probability distribution
size_distribution <- restaurant_groups %>%
  count(group_size) %>%
  mutate(probability = n / sum(n))

# Calculate probability of picking group of 4 or more
size_distribution %>%
  # Filter for groups of 4 or larger
  filter(group_size >= 4) %>%
  # Calculate prob_4_or_more by taking sum of probabilities
  summarize(prob_4_or_more = sum(probability))
```

Since the sample mean will likely be closer to 5 (the expected value) with larger sample sizes, you have a better chance of getting a sample mean further away from 5 with a smaller sample.

# Continuous Distribution

## Uniform distribution

```{r}
# Probability of wait time <= 7
punif(7, min = 0, max = 12)
```

```{r}
# Probability of wait time >= 7
punif(7, min = 0, max = 12, lower.tail = FALSE)
```

- $P(4 \leq \text{wait time} \leq 7 )$

```{r}
punif(7, min = 0, max = 12) - punif(4, min = 0, max = 12)
```

## Data back-ups

The sales software used at your company is set to automatically back itself up, but no one knows exactly what time the back-ups happen. It is known, however, that back-ups happen exactly every 30 minutes. Amir comes back from sales meetings at random times to update the data on the client he just met with. He wants to know how long he'll have to wait for his newly-entered data to get backed up. Use your new knowledge of continuous uniform distributions to model this situation and answer Amir's questions.

To model how long Amir will wait for a back-up using a continuous uniform distribution, save his lowest possible wait time as min and his longest possible wait time as max. Remember that back-ups happen every 30 minutes.

```{r}
# Min and max wait times for back-up that happens every 30 min
min <- 0
max <- 30
```

Calculate the probability that Amir has to wait less than 5 minutes, and store in a variable called prob_less_than_5.

```{r}
# Calculate probability of waiting less than 5 mins
prob_less_than_5 <- punif(5, min, max)
prob_less_than_5
```

Calculate the probability that Amir has to wait more than 5 minutes, and store in a variable called prob_greater_than_5.

```{r}
# Calculate probability of waiting more than 5 mins
prob_greater_than_5 <- punif(5, min, max, lower.tail = FALSE)
prob_greater_than_5
```

Calculate the probability that Amir has to wait between 10 and 20 minutes, and store in a variable called prob_between_10_and_20.

```{r}
# Calculate probability of waiting 10-20 mins
prob_between_10_and_20 <- punif(20, min, max) - punif(10, min, max)
prob_between_10_and_20
```

# The binomial distribution

```{r}
simulation_nb <- seq(1, 1000)
```

```{r}
set.seed(334)
```

Generate 1000 wait times from the continuous uniform distribution that models Amir's wait time. Add this as a new column called time in the wait_times data frame.
```{r}
time <- runif(1000, min, max)
```

```{r}
wait_times <- tibble(simulation_nb = simulation_nb,
                     time = time)
```

Create a histogram of the simulated wait times with 30 bins.

```{r}
# Generate 1000 wait times between 0 and 30 mins, save in time column
wait_times %>%
  mutate(time = runif(1000, min = 0, max = 30)) %>%
  # Create a histogram of simulated times
  ggplot(aes(time)) +
  geom_histogram(bins = 30)

```

