---
title: "Introduction do Data Visualization with ggplot2"
author: "by Edneide Ramalho"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
      highlight: textmate
      logo: logo.png
      theme: flatly
      number_sections: yes
      toc: yes
      toc_float:
        collapsed: yes
        smooth_scroll: no
---

# Drawing your first plot

To get a first feel for `ggplot2`, let's try to run some basic `ggplot2` commands. The `mtcars` dataset contains information on 32 cars from a 1973 issue of Motor Trend magazine. This dataset is small, intuitive, and contains a variety of continuous and categorical variables.

-   Load the `ggplot2` package using [**`library()`**](http://www.rdocumentation.org/packages/base/functions/library).

-   Use [**`str()`**](http://www.rdocumentation.org/packages/utils/functions/str) to explore the structure of the `mtcars` dataset.

-   Hit *Submit Answer*. This will execute the example code on the right. See if you can understand what `ggplot` does with the data.

```{r}
# Load the ggplot2 package
library(ggplot2)

# Explore the mtcars data frame with str()
str(mtcars)

# Execute the following command
ggplot(mtcars, aes(cyl, mpg)) +
  geom_point()
```

## Data columns types affect plot types

The plot from the previous exercise wasn't really satisfying. Although `cyl` (the number of cylinders) is categorical, you probably noticed that it is classified as numeric in `mtcars`. This is really misleading because the representation in the plot doesn't match the actual data type. You'll have to explicitly tell `ggplot2` that `cyl` is a categorical variable.

-   Change the [**`ggplot()`**](http://www.rdocumentation.org/packages/ggplot2/functions/ggplot) command by wrapping [**`factor()`**](http://www.rdocumentation.org/packages/base/functions/factor) around `cyl`.

-   Hit *Submit Answer* and see if the resulting plot is better this time.

```{r}
# Load the ggplot2 package
library(ggplot2)

# Change the command below so that cyl is treated as factor
ggplot(mtcars, aes(factor(cyl), mpg)) +
  geom_point()
```

# Mapping data columns to aesthetics

Let's dive a little deeper into the three main topics in this course: The data, aesthetics, and geom layers. We'll get to making pretty plots in the last chapter with the themes layer.

We'll continue working on the 32 cars in the `mtcars` data frame.

Consider how the examples and concepts we discuss throughout these courses apply to your own data-sets!

-   Add a color aesthetic mapped to the displacement of the car engine: inside [**`aes()`**](http://www.rdocumentation.org/packages/ggplot2/functions/aes), add a `color` argument equal to `disp`.

```{r}
# Edit to add a color aesthetic mapped to disp
ggplot(mtcars, aes(wt, mpg, color = disp))+
  geom_point()
```

-   This time, map `disp` to the `size` aesthetic.

```{r}
# Change the color aesthetic to a size aesthetic
ggplot(mtcars, aes(wt, mpg, size = disp)) +
  geom_point()
```

## Understanding variables

In the previous exercise you saw that `disp` can be mapped onto a color gradient or onto a continuous size scale.

Another argument of [**`aes()`**](http://www.rdocumentation.org/packages/ggplot2/functions/aes) is the `shape` of the points. There are a finite number of shapes which [**`ggplot()`**](http://www.rdocumentation.org/packages/ggplot2/functions/ggplot) can automatically assign to the points. However, if you try this command in the console to the right:

    ggplot(mtcars, aes(wt, mpg, shape = disp)) +
      geom_point()

It gives an error. What does this mean?

-   **Erro**: A continuous variable can not be mapped to shape

# ggplot2 layers

```{r}
g <- ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_jitter(alpha = 0.5)
g
```

```{r}
g +
  theme_classic()
```

## Adding geometries

The `diamonds` dataset contains details of 1,000 diamonds. Among the variables included are `carat` (a measurement of the diamond's size) and `price`.

You'll use two common geom layer functions:

-   [**`geom_point()`**](http://www.rdocumentation.org/packages/ggplot2/functions/geom_point) adds points (as in a scatter plot).

-   [**`geom_smooth()`**](http://www.rdocumentation.org/packages/ggplot2/functions/geom_smooth) adds a smooth trend curve.

As you saw previously, these are added using the `+` operator.

    ggplot(data, aes(x, y)) +
      geom_*()

Where `*` is the specific geometry needed.

-   Explore the `diamonds` data frame with the [**`str()`**](http://www.rdocumentation.org/packages/utils/functions/str) function.

```{r}
str(diamonds)
```

-   Edit the plot code to add a point geom. Use the `+` operator to add [**`geom_point()`**](http://www.rdocumentation.org/packages/ggplot2/functions/geom_point) to the [**`ggplot()`**](http://www.rdocumentation.org/packages/ggplot2/functions/ggplot) command.

```{r}
# Add geom_point() with +
ggplot(diamonds, aes(carat, price)) +
  geom_point()
```

-   Add a smooth geom to the plot. Use the `+` operator to add [**`geom_smooth()`**](http://www.rdocumentation.org/packages/ggplot2/functions/geom_smooth).

```{r}
# Add geom_point() with +
ggplot(diamonds, aes(carat, price)) +
  geom_point() + 
  geom_smooth()
```

## Changing one geom or every geom

If you have multiple geoms, then mapping an aesthetic to data variable inside the call to `ggplot()` will change *all* the geoms. It is also possible to make changes to *individual* geoms by passing arguments to the `geom_*()` functions.

`geom_point()` has an `alpha` argument that controls the opacity of the points. A value of `1` (the default) means that the points are totally opaque; a value of `0` means the points are totally transparent (and therefore invisible). Values in between specify transparency.

The plot you drew last time is provided in the script.

-   Edit the plot code to map the `color` aesthetic to the `clarity` data variable.

```{r}
# Map the color aesthetic to clarity
ggplot(diamonds, aes(carat, price, color = clarity)) +
  geom_point() +
  geom_smooth()
```

```{r}
# Make the points 40% opaque
ggplot(diamonds, aes(carat, price, color = clarity)) +
  geom_point(alpha = 0.4) +
  geom_smooth()
```

## Saving plots as variables
