---
title: "Visible aesthetics"
author: "by Edneide Ramalho"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
      highlight: textmate
      logo: logo.png
      theme: flatly
      number_sections: yes
      toc: yes
      toc_float:
        collapsed: yes
        smooth_scroll: no
---

# Mapping onto the color aesthetic

```{r packages}
library(tidyverse)
```

```{r}
iris %>% 
  ggplot(aes(x = Sepal.Length,
             y = Sepal.Width,
             color = Species)) +
  geom_point()
```

## Exercises

### **All about aesthetics: color, shape and size**

In the video you saw 9 visible aesthetics. Let's apply them to a categorical variable --- the cylinders in `mtcars`, `cyl`.

These are the aesthetics you can consider within [**`aes()`**](http://www.rdocumentation.org/packages/ggplot2/functions/aes) in this chapter: `x`, `y`, `color`, `fill`, `size`, `alpha`, `labels` and `shape`.

One common convention is that you don't name the `x` and `y` arguments to `aes()`, since they almost always come first, but you do name other arguments.

In the following exercise the `fcyl` column is categorical. It is `cyl` transformed into a `factor`.

1.  Map `mpg` onto the `x` aesthetic, and `fcyl` onto the `y`.

```{r}
mtcars_new <- mtcars %>% 
  dplyr::mutate(fcyl = factor(cyl))
```

```{r}
# Map x to mpg and y to fcyl
ggplot(mtcars_new, aes(mpg, fcyl)) +
  geom_point()
```

2.  Swap the mappings of the first plot: `fcyl` onto the `x` aesthetic, and `mpg` onto the `y`.

```{r}
# Swap mpg and fcyl
ggplot(mtcars_new, aes(fcyl, mpg)) +
  geom_point()
```

3.  Map `wt` onto `x`, `mpg` onto `y`, and `fcyl` onto `color`.

```{r}
# Map x to wt, y to mpg and color to fcyl
ggplot(mtcars_new, aes(wt, mpg, color = fcyl)) +
  geom_point()
```

4.  Modify the point layer of the previous plot by changing the `shape` argument to `1` and increasing the `size` to `4`.

```{r}
ggplot(mtcars_new, aes(wt, mpg, color = fcyl)) +
  # Set the shape and size of the points
  geom_point(shape = 1, size = 4)
```
