---
title: "Dimension Reduction"
author: "by Edneide Ramalho"
date: "02/08/2021"
output: html_document
---

In this lesson we'll discuss principal component analysis (PCA) and singular
 value decomposition (SVD), two important and related techniques of dimension
 reduction. This last entails processes which finding subsets of variables in
 datasets that contain their essences. PCA and SVD are used in both the
 exploratory phase and the more formal modelling stage of analysis. We'll focus
 on the exploratory phase and briefly touch on some of the underlying theory.

```{r}
library(readr)
dataMatrix <- read_csv("dataMatrix.csv")
head(dataMatrix)
```

```{r}
dataMatrix <- dataMatrix[, -1]
tempMatrix <- as.matrix(dataMatrix)
heatmap(tempMatrix)
```

We can see that even with the clustering that heatmap provides, permuting the
 rows (observations) and columns (variables) independently, the data still looks
 random.

Let's add a pattern to the data. We've put some R code in the file addPatt.R for
 you. Run the command myedit with the single argument "addPatt.R" (make sure to
 use the quotation marks) to see the code. You might have to click your cursor in
 the console after you do this to keep from accidentally changing the file.

So in rows affected by the coin flip, the 5 left columns will still have a mean
\| of 0 but the right 5 columns will have a mean closer to 3.

```{r}
addPatt <- function(){
  set.seed(678910)
for(i in 1:40){
  # flip a coin
  coinFlip <- rbinom(1,size=1,prob=0.5)
  # if coin is heads add a common pattern to that row
  if(coinFlip){
    dataMatrix[i,] <- dataMatrix[i,] + rep(c(0,3),each=5)
  }
}
  return(dataMatrix)
}
```

```{r}
new_dataMatrix <- addPatt()
```

```{r}
heatmap(as.matrix(new_dataMatrix))
```

Here's the image of the altered dataMatrix after the pattern has been added. The
\| pattern is clearly visible in the columns of the matrix. The right half is
\| yellower or hotter, indicating higher values in the matrix.

Again we see the pattern in the columns of the matrix. As shown in the
 dendrogram at the top of the display, these split into 2 clusters, the lower
 numbered columns (1 through 5) and the higher numbered ones (6 through 10).
 Recall from the code in addPatt.R that for rows selected by the coinflip the
 last 5 columns had 3 added to them. The rows still look random.
