---
title: "Week 4"
author: "Edneide Ramalho"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
      highlight: textmate
      logo: logo.png
      theme: cerulean
      number_sections: yes
      toc: yes
      toc_float:
        collapsed: yes
        smooth_scroll: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data

-   Train data set

```{r}
library(readr)
X_train <- read_table2("Week 4/UCI HAR Dataset/train/X_train.txt", 
    col_names = FALSE)
dim(X_train)
```

-   Training labels:

1 WALKING

2 WALKING_UPSTAIRS

3 WALKING_DOWNSTAIRS

4 SITTING

5 STANDING

6 LAYING

```{r}
library(readr)
y_train <- read_csv("Week 4/UCI HAR Dataset/train/y_train.txt", 
    col_names = FALSE)
```

-   Mudando o nome da variável:

```{r}
names(y_train) <- "activity"
```

```{r}
table(y_train$activity)
```

-   Adding the activity name to `y_train`:

```{r}
library(readr)
activity_labels <- read_table2("Week 4/UCI HAR Dataset/activity_labels.txt", 
    col_names = FALSE)
names(activity_labels)
```

-   Join `activity_labels` to `y_train`:

```{r}
library(tidyverse)
y_train <- y_train %>% 
  left_join(activity_labels, by = c("activity" = "X1"))
glimpse(y_train)
names(y_train)[2]  <- "activity_name"
```

```{r}
glimpse(y_train)
```

-   Importing features:

    ```{r}
    library(readr)
    features <- read_table2("Week 4/UCI HAR Dataset/features.txt", 
        col_names = FALSE)
    names(features)[2] <- "features"
    ```

-   Replacing features names to the `X_train` dataset and adding the activity label.

```{r}
names(X_train) <- features$features
```

```{r}
samsung_data <- cbind(y_train, X_train)
```

-   We'd like to separate the activities depending on the values we have for the features.

```{r}
table(samsung_data$activity_name)
```

-   Adding the subject name:

```{r}
subject_train <- read_csv("Week 4/UCI HAR Dataset/train/subject_train.txt", col_names = FALSE)
names(subject_train) <- "subject"
```

```{r}
samsung_data <- cbind(subject_train, samsung_data)
names(samsung_data)[1:6]
```

# Plotting average accelaration for the first subject

```{r, include=FALSE}
par(mfrow = c(1, 2), mar = c(5, 4, 1, 1))
samsung_data <- transform(samsung_data, activity_name = factor(activity_name))
sub1 <- subset(samsung_data, subject == 1)
plot(sub1[, 4], col = sub1$activity_name, ylab = names(sub1[4]))
plot(sub1[, 5], col = sub1$activity_name, ylab = names(sub1[5]))
legend("right", 
       legend = unique(sub1$activity_name, 
                       col = unique(sub1$activity_name), 
                       pch = 1))
```

```{r}
library(patchwork)
library(ggplot2)
library(magrittr)
library(tidyverse)

df1 <- samsung_data %>% 
  filter(subject == 1)
df1$index <- 1:dim(df1)[1]

plot1 <- ggplot(df1, aes(x = index, y = tBodyAcc.mean...X, color = activity_name)) +
  geom_point() +
  theme(legend.position = "bottom",
        legend.title = element_blank())

plot2 <- ggplot(df1, aes(x = index, y = tBodyAcc.mean...Y, color = activity_name)) +
  geom_point() +
  theme(legend.position = "none",
        legend.title = element_blank())

plot1 + plot2
```

# Clustering based just on average acceleration

```{r}
# library(devtools) # get from CRAN with install.packages("devtools")
# install_github("ririzarr/rafalib")

library(rafalib)
distance_matrix <- dist(sub1[, 4:6])
hclustering <- hclust(distance_matrix)
par(mfrow = c(1, 1), mar = c(5, 4, 1, 1))
myplclust(hclustering, lab.col = unclass(sub1$activity_name))
```

The colors for specific activities are still messed up. We cannot see a clear pattern on the dendrogram.

# Plotting max acceleration for the first subject

```{r}
names(sub1)[1:20]
```

```{r}
par(mfrow = c(1, 2))
plot(sub1[, 13], pch = 19, 
     col = sub1$activity_name, ylab = names(sub1)[13])
plot(sub1[, 14], pch = 19, 
     col = sub1$activity_name, ylab = names(sub1)[14])
```

The maximum accelaration seems to separate more clearly the moving and not moving activities.

Now, let's cluster based on the maximum accelaration

```{r}
distance_matrix2 <- dist(sub1[, 13:15])
hclustering2 <- hclust(distance_matrix2)
par(mfrow = c(1, 1))
myplclust(hclustering2,
          lab.col = unclass(sub1$activity_name))
```

Now we can see on the left-hand side walking activities and in the right-hand side we can see the not-moving activities.

Maximum accelaration seem to cluster well moving from not-moving activities. But inside of these two big clusters, it's difficult to see the division of the activities inside them.

# Singular Value Decomposition

```{r}
svd1 = svd(scale(sub1[, -c(1, 2, 3)]))
par(mfrow = c(1, 2))
plot(svd1$u[, 1], col = sub1$activity_name, pch = 19)
plot(svd1$u[, 2], col = sub1$activity_name, pch = 19)
```

The first singular vector seem to separate quite well the moving from the non-moving activities. The second singular value seems to separate well the magenta color from the others.

# New clustering with maximum contributer

```{r}
max_contrib <- which.max(svd1$v[, 2])
distance_matrix3 <- dist(sub1[, c(13:15, max_contrib)])
hclustering3 <- hclust(distance_matrix3)
par(mfrow = c(1, 1))
myplclust(hclustering3, lab.col = unclass(sub1$activity_name))
```

-   The variable with the maximum contribution is:

```{r}
names(samsung_data)[max_contrib]
```

# K-means clustering (nstart = 1, first try)

```{r}
k_clust <- kmeans(sub1[, -c(1, 2, 3)], center = 6)
table(k_clust$cluster, sub1$activity_name)
```

# K-means clustering (nstart = 1, first try)

```{r}
k_clust <- kmeans(sub1[, -c(1, 2, 3)], center = 6, nstart = 100)
table(k_clust$cluster, sub1$activity_name)
```

-   Cluster 1: mix of laying, sitting and standing.

-   Cluster 2: walking

-   Cluster 3: sitting and standing

-   Cluster 4: walking down

-   Cluster 5: laying and walking up

-   Cluster 6: laying

**Why does the K-means algorithm produce different clustering solutions every time you run it?**

K-Means chooses are random starting point by default.

# Cluster 1: Variable Centers (Laying)

```{r}
plot(k_clust$centers[1, 1:10], pch = 19, 
     ylab = "Cluster Center",
     xlab = "")
```

We can see what features can contribute more to predict the activity.
